---
title: "PS270 Final Project"
author: "Juliet Chin"
description: My final project
output:
  html_document:
    self_contained: true
  pdf_document: default
---
```{r setup, include=FALSE}
library(tidyverse)
library(knitr)

starwars <- read_csv("StarWars.csv", show_col_types = FALSE)
```

## Introduction

My research question asks whether gender predicts whether someone considers themselves a Star Wars fan. I hypothesize that men are more likely than women to identify as Star Wars fans. One theoretical explanation for this expectation is that science fiction media has historically been marketed more heavily toward men. If certain genres are framed as masculine-coded interests, men may be more likely to internalize fandom as part of their identity. As social scientists, we are interested in whether this broader pattern of gendered media consumption appears systematically in survey data.
 
## Data

This project uses survey data from FiveThirtyEight’s Star Wars dataset, which contains respondents’ views on Star Wars fandom and demographic characteristics such as gender. Because this dataset surveys individuals at one point in time, this research design is cross-sectional. This means I can examine whether gender is associated with Star Wars fandom, though I cannot determine causality.

My dependent variable is whether a respondent identifies as a Star Wars fan (Yes = 1, No = 0). My independent variable is gender, measured as Male or Female based on survey responses. I hypothesize that men will be more likely than women to identify as Star Wars fans. My hypothesis will be supported if men show a higher predicted probability of fandom and if the gender coefficient in my regression analysis is statistically significant.

```{r outcome-plot, message=FALSE, warning=FALSE}
fan_gender <- starwars |>
  select(
    fan_response = `Do you consider yourself to be a fan of the Star Wars film franchise?`,
    gender = Gender
  ) |>
  filter(!is.na(fan_response), !is.na(gender)) |>
  filter(gender %in% c("Male", "Female")) |>
  mutate(fan = if_else(fan_response == "Yes", 1, 0)) |>
  group_by(gender) |>
  summarize(fan_percent = mean(fan) * 100, .groups = "drop")

ggplot(fan_gender, aes(x = gender, y = fan_percent, fill = gender)) +
  geom_col(width = 0.6) +
  scale_fill_manual(
    values = c("Male" = "lightblue", "Female" = "pink"),
    guide = "none"
  ) +
  labs(
    title = "Star Wars Fandom by Gender",
    x = "Gender",
    y = "Percent Who Identify as a Star Wars Fan"
  ) +
  ylim(0, 100) +
  theme_minimal()
```


Write up: 
This plot shows the main outcome variable, Star Wars fandom, by gender. About 60 percent of women in the cleaned sample identify as Star Wars fans, compared to about 72 percent of men. This pattern supports my hypothesis descriptively because men report higher levels of Star Wars fandom than women.

## Results

```{r regression-table, message=FALSE, warning=FALSE}
fan_model_data <- starwars |>
  select(
    fan_response = `Do you consider yourself to be a fan of the Star Wars film franchise?`,
    gender = Gender
  ) |>
  filter(!is.na(fan_response), !is.na(gender)) |>
  filter(gender %in% c("Male", "Female")) |>
  mutate(
    fan = if_else(fan_response == "Yes", 1, 0),
    male = if_else(gender == "Male", 1, 0)
  )

model <- glm(fan ~ male, data = fan_model_data, family = binomial)

regression_table <- tibble(
  Term = c("Intercept", "Male"),
  Estimate = round(coef(summary(model))[, "Estimate"], 3),
  `Standard Error` = round(coef(summary(model))[, "Std. Error"], 3),
  `P-value` = if_else(
  coef(summary(model))[, "Pr(>|z|)"] < 0.001,
  "<0.001",
  as.character(round(coef(summary(model))[, "Pr(>|z|)"], 3))
)
)

kable(
  regression_table,
  caption = "Logistic Regression Predicting Star Wars Fandom"
)
```

To test whether gender predicts Star Wars fandom, I ran a logistic regression with fan identity (1 = fan, 0 = not a fan) as the outcome and gender (1 = Male, 0 = Female) as the predictor. The results support my hypothesis: being male is a statistically significant positive predictor of Star Wars fandom (B = 0.52, p < 0.001). Men have approximately 1.69 times higher odds of identifying as a Star Wars fan compared to women. This is consistent with the idea that science fiction media has historically been marketed more toward men, shaping gendered patterns of fandom identity. However, I do not interpret this relationship as causal because this is cross-sectional survey data and the model does not account for other possible factors, such as age, childhood media exposure, or broader media interests.

## Conclusion
Overall, my findings support my hypothesis that men are more likely than women to identify as Star Wars fans. Both the descriptive bar plot and logistic regression show that men report significantly higher levels of Star Wars fandom than women. However, because this analysis uses cross-sectional survey data and only includes gender as a predictor, I cannot conclude that gender itself causes fandom differences. Other factors, such as age, socialization, or broader media exposure, may also shape this relationship. Future research could improve this analysis by including additional variables and testing whether gender remains significant after accounting for these potential confounding factors.













